---
title: "K-means clustering on Air Pollution"
author: "Neha"
date: "3/4/2017"
output: html_document
---
The data produced by IoT (internet of things) is enormous and data mining techniques can be used to get hidden information, which is of high business value. Smart cities are completely based on IoT.Air pollution is increasing rapidly in the smart cities and has adverse effects on human health. The sources of pollution are many including
road traffic, industrial gases and others.
In this study we try to find the healthiest areas, which are suitable for leaving, in the smart cities by using K-means clustering. The dataset is generated from the City Plus project. The data is enormous and dynamic due to the number of sensors deployed in the same location and their measurement frequency.
This data consists of 5 air pollutants namely ozone, sulfur dioxide, nitrogen dioxide, carbon monoxide and particulate matter. There are 3 more fields in the data set namely- Longitude, latitude and timestamp.

```{r include=FALSE, cache=FALSE}
library(readr)
library(dplyr)
library(data.table)
library(sqldf)
library(ggplot2)
library(ggmap)
library(maps)
```


Get the data
```{r, include=FALSE}
setwd("/Users/neha/Documents/Capstone/pollution")
temp = list.files(pattern="*.csv")
tbl = lapply(temp, read_csv) %>% bind_rows()
```



Get the ozone concentration grouped by latitude and longitude
```{r}
x1 = sqldf("select longitude, latitude, avg(ozone) as avg_ozone, avg(particullate_matter) as particulate from tbl group by longitude,latitude")
head(x1)
```

### K-means
Here we apply K-means algorithm for k = 4 through 10. The objective is to group the data into clusters and identify the cleanest areas in the city based on ozone level concentration.
```{r}
# k = 4
result4<-kmeans(x1,4)
result4$centers
```

```{r}
# k = 5
result5<-kmeans(x1,5)
result4$centers
```

```{r}
# k = 6
result6<-kmeans(x1,6)
result6$centers
```

```{r}
# k = 7
result7<-kmeans(x1,7)
result7$centers
```

```{r}
# k = 8
result8<-kmeans(x1,8)
result8$centers
```

```{r}
# k = 9
result9<-kmeans(x1,9)
result9$centers
```

```{r}
# k = 10
result10<-kmeans(x1,10)
result10$centers
```

#### Visualizing the cluster
```{r, echo=FALSE}
plot(x1[c("avg_ozone","particulate")], col = result10$cluster )
```

#### Mapping the cluster
```{r, echo=FALSE}
df = result10$centers
df<-as.data.frame(df)
min_ozone_level = min(df[,3]) # get min level of ozone
lon <- as.vector(df['longitude'])
lat <- as.vector(df['latitude'])
df1 <- as.data.frame(cbind(lon,lat))
newdata <- subset(df, avg_ozone== min_ozone_level, 
select=c(longitude, latitude))
long <-as.vector(newdata['longitude'])
lati <-as.vector(newdata['latitude'])

mapgilbert <- get_map(location = c(lon = mean(df1$lon), lat = mean(df1$lat)), zoom = 12, maptype = "satellite", scale = 1)
# plotting the map with some points on it
ggmap(mapgilbert) + geom_point(data = df1, aes(x = lon, y = lat, fill = "red", alpha = 0.1), size = 3, shape = 21) + guides(fill=FALSE, alpha=FALSE, size=FALSE)

#mapgilbert <- get_map(location = c(lon = mean(df1$lon), lat = mean(df1$lat)), zoom = 12, maptype = "satellite", scale = 1)
# plotting the map with some points on it
#ggmap(mapgilbert) + geom_point(data = newdata, aes(x = long, y = lati, fill = "red", alpha = 0.1), size = 5, shape = 21) + guides(fill=FALSE, alpha=FALSE, size=FALSE)
```

###Conclusion
When K = 10, we find the ozone level concentration to be the lowest. Thus the cleanest
area is found to be at:
Longitude:
```{r}
long

```

Latitude:
```{r}
lati

```

Minimum ozone level:
```{r}
min_ozone_level
```



